### 2024-04-18 paper/paper.qmd
model data added and other changes made

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - First author
  - Another author
thanks: "Code and data are available at: LINK."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(ggplot2)
library(modelsummary)

```

# Introduction

You can and should cross-reference sections and sub-sections. We use @citeR and @rohan.

The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}
#| label: hist-hap
#| fig-cap: Distribution of Happiness Scores
#| echo: false
#| warning: false
#| message: false

analysis_data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))
analysis_data$IHDI <- round(as.numeric(analysis_data$IHDI), 3)

# View the changes to ensure they have been applied
head(analysis_data$IHDI)

ggplot(analysis_data, aes(x = Happiness_score)) +
  geom_histogram(binwidth = 0.2, fill = "skyblue", color = "black") +
  labs(x = "Happiness Score", y = "Frequency") +
  theme_minimal()
```

```{r}
#| label: hist-gini
#| fig-cap: Distribution of Gini Index
#| echo: false
#| warning: false
#| message: false

# Histogram of Gini Coefficients
ggplot(analysis_data, aes(x = Gini_Index)) +
  geom_histogram(binwidth = 1, fill = "lightgreen", color = "black") +
  labs(x = "Gini index", y = "Frequency") +
  theme_minimal()

```

```{r}
#| label: scat-hap-gini
#| fig-cap: Relationship between Gini Index and Happiness
#| echo: false

# Load the necessary libraries
library(ggplot2)

# Assuming analysis_data is your data frame and it has 'Gini_Coefficient' and 'Happiness_Score' columns
# Basic scatter plot
basic_plot <- ggplot(analysis_data, aes(x = Gini_Index, y = Happiness_score)) +
  geom_point()

# Display the plot
print(basic_plot)

```

```{r}
#| label: scat-hap-ihdi
#| fig-cap: Relationship between IHDI and Happiness
#| echo: false

# Scatter plot between IHDI and Happiness Score
# Load the necessary library
library(ggplot2)

# Assuming analysis_data is your data frame and it has 'IHDI' and 'Happiness_Score' columns
# Basic scatter plot for IHDI vs. Happiness Score
basic_plot_ihdi <- ggplot(analysis_data, aes(x = IHDI, y = Happiness_score)) +
  geom_point()

# Display the plot
print(basic_plot_ihdi)

```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

Talk way more about it.

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

```{=tex}
\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}
```
We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

first_model <-
  readRDS(file = here::here("models/first_model.rds"))

second_model <-
  readRDS(file = here::here("models/second_model.rds"))

```

```{r}
#| echo: false
#| eval: true
#| label: tbl-secondtmodelresults
#| tbl-cap: "Multiple Linear Regression"
#| warning: false

# Generate the model summary for 'first_model' with the desired statistics
modelsummary::modelsummary(
  list(
    "Second model" = second_model
  ),
  statistic = c("std.error", "statistic", "p.value", "conf.int")
)
```

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {.unnumbered}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

```

\newpage

# References