### 2024-04-18 paper/paper.qmd
abstract

---
title: "My title"
author: 
  - Carl Fernandes
  - Another author
thanks: "Code and data are available at: LINK."
date: today
date-format: long
abstract: "This study investigates the connection between income inequality and national hapiness, questioning the impact of the Gini index. By examining cross-sectional data we find that income inequality has a limited direct effect on happiness. It is evident that factors like the quality of institutions and levels of development as measured by the Democracy Index and the Inequality adjusted Human Development Index (IHDI) have a more significant impact on shaping happiness. Our results indicate that focusing solely on income inequality doesn't provide a complete understanding. Instead a wider range of indicators are. This research contributes to the conversation on disparity and satisfaction by emphasizing the importance of using multifaceted approaches in happiness research and policy development.
"
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(ggplot2)
library(modelsummary)

```

# Introduction

You can and should cross-reference sections and sub-sections. We use @citeR and @rohan.

The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}
#| label: hist-hap
#| fig-cap: Distribution of Happiness Scores
#| echo: false
#| warning: false
#| message: false

analysis_data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))
analysis_data$IHDI <- round(as.numeric(analysis_data$IHDI), 3)

# View the changes to ensure they have been applied
head(analysis_data$IHDI)

ggplot(analysis_data, aes(x = Happiness_score)) +
  geom_histogram(binwidth = 0.2, fill = "skyblue", color = "black") +
  labs(x = "Happiness Score", y = "Frequency") +
  theme_minimal()
```

```{r}
#| label: hist-gini
#| fig-cap: Distribution of Gini Index
#| echo: false
#| warning: false
#| message: false

# Histogram of Gini Coefficients
ggplot(analysis_data, aes(x = Gini_Index)) +
  geom_histogram(binwidth = 1, fill = "lightgreen", color = "black") +
  labs(x = "Gini index", y = "Frequency") +
  theme_minimal()

```

```{r}
#| label: scat-hap-gini
#| fig-cap: Relationship between Gini Index and Happiness
#| echo: false

# Load the necessary libraries
library(ggplot2)

# Assuming analysis_data is your data frame and it has 'Gini_Coefficient' and 'Happiness_Score' columns
# Basic scatter plot
basic_plot <- ggplot(analysis_data, aes(x = Gini_Index, y = Happiness_score)) +
  geom_point()

# Display the plot
print(basic_plot)

```

```{r}
#| label: scat-hap-ihdi
#| fig-cap: Relationship between IHDI and Happiness
#| echo: false

# Scatter plot between IHDI and Happiness Score
# Load the necessary library
library(ggplot2)

# Assuming analysis_data is your data frame and it has 'IHDI' and 'Happiness_Score' columns
# Basic scatter plot for IHDI vs. Happiness Score
basic_plot_ihdi <- ggplot(analysis_data, aes(x = IHDI, y = Happiness_score)) +
  geom_point()

# Display the plot
print(basic_plot_ihdi)
```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

Talk way more about it.

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

```{=tex}
\begin{align} 
\text{HappinessScore}_i &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta \times \text{GiniIndex}_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}
```

```{=tex}
\begin{align} 
\text{Happiness\_score}_i &= \beta_0 + \beta_1 \times \text{Democracy\_index}_i + \beta_2 \times \text{Gini\_index}_i + \beta_3 \times \text{IHDI}_i + \varepsilon_i \\
\varepsilon_i &\sim \mbox{Normal}(0, \sigma^2)
\end{align}
```
We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

first_model <-
  readRDS(file = here::here("models/first_model.rds"))

second_model <-
  readRDS(file = here::here("models/second_model.rds"))

```

```{r}
#| echo: false
#| eval: true
#| label: tbl-firsttmodelresults
#| tbl-cap: "Linear Regression"
#| warning: false

# Generate the model summary for 'first_model' with the desired statistics
modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  statistic = c("std.error", "statistic", "p.value", "conf.int")
)
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-secondtmodelresults
#| tbl-cap: "Multiple Linear Regression"
#| warning: false

# Generate the model summary for 'first_model' with the desired statistics
modelsummary::modelsummary(
  list(
    "Second model" = second_model
  ),
  statistic = c("std.error", "statistic", "p.value", "conf.int")
)
```

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {.unnumbered}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

```

\newpage

# References