### 2024-04-19 other/datasheet/datasheet_template.qmd
extra stuff

---
title: "Datasheet for 'Income Inequality and Happiness"
author: 
  - Carl Fernandes
thanks: "Code and data are available at: https://github.com/Carl-1506/Happiness-and-inequality.git"
date: "`today"
date-format: "D MMMM YYYY"
abstract: "This study examines the relationship between income inequality, as measured by the Gini Index, and national happiness scores. Incorporating other socio-economic factors such as the Democracy Index and Inequality-Adjusted Human Development Index, we explore the depth of this relationship. The study aims to inform policy by identifying the broader determinants of happiness beyond economic measures."
format: pdf
number-sections: true
editor: 
  markdown: 
    wrap: sentence
---

Extract of the questions from @gebru2021datasheets.

**Motivation**

1.  *For what purpose was the dataset created? Was there a specific task in mind? Was there a specific gap that needed to be filled? Please provide a description.*
    -   The dataset was curated to explore the hypothesis that income inequality, as quantified by the Gini Index, has a discernible impact on the perceived happiness within nations, taking into account factors like democratic governance and human development.
2.  *Who created the dataset (for example, which team, research group) and on behalf of which entity (for example, company, institution, organization)?*
    -   Dataset was created by Carl Fernandes for University of Toronto
3.  *Who funded the creation of the dataset? If there is an associated grant, please provide the name of the grantor and the grant name and number.*
    -   N/A
4.  *Any other comments?*
    -   N/A

**Composition**

1.  **What do the instances that comprise the dataset represent?**

    The dataset instances represent individual countries, with each instance encapsulating various socio-economic indicators for the year 2022.

2.  **How many instances are there in total?**

    There are 100 instances, each corresponding to a unique country.

3.  **Does the dataset contain all possible instances or is it a sample of instances from a larger set?**

    The dataset is a sample of countries from the year 2022, selected based on the availability of complete data for the variables of interest.

4.  **What data does each instance consist of?**

    Each instance consists of the country's name, the Gini Index, Democracy Index, Inequality-Adjusted Human Development Index (IHDI), and the corresponding Happiness Score for 2022.

5.  **Collection process**

6.**How was the data associated with each instance acquired?**
The data was acquired from publicly available databases, including the World Happiness Report, the World Bank, Our World in Data, and the UNDP.

7. **What mechanisms or procedures were used to collect the data?**
Data were collected through automated scripts pulling from the aforementioned databases' APIs or data download sections.

**Preprocessing/cleaning/labeling**

1.  **Was any preprocessing/cleaning/labeling of the data done?**
Preprocessing involved cleaning for missing values, ensuring consistency across indicators, and standardizing country names.

**Uses**

1.  **Has the dataset been used for any tasks already?**
The dataset has been used for an initial exploratory data analysis and regression modeling to ascertain the relationships between happiness and the selected socio-economic indicators.

**Distribution**

1.  **How will the dataset be distributed?**
The dataset, along with the code, will be available in a GitHub repository, with a README explaining the structure and use-cases.

**Maintenance**

1.  *Who will be supporting/hosting/maintaining the dataset?*
    -   Github

\newpage